---
title: "DSC 441 Homework 5"
author: "Vijayalakshmi Ramesh"
date: "`r Sys.Date()`"
output: word_document
---

Credit Card Customers Prediction

#About data:

**Problem:** 
A manager at the bank is disturbed with more and more customers leaving their credit card services. They would really appreciate if one could predict for them who is gonna get churned so they can proactively go to the customer to provide them better services and turn customers' decisions in the opposite direction.

A business manager of a consumer credit card portfolio is facing the problem of customer attrition. They want to analyze the data to find out the reason behind this and leverage the same to predict customers who are likely to drop off.

Attrited customers are the group of customers that has been lost through time for any reason. The process of losing a client is referred to as 
customer attrition, also known as customer churn.

**Goal:** To predict churning customers

**Details of dataset:**

```{r}
bank_churners <- read.csv("Bank_Churners.csv")

library(tidyverse)
bank_churners <- bank_churners %>% select(-c(1,10,22,23)) 
str(bank_churners)
```

- There are 23 variables out of that 2 are from Naive Bayes, so I'll omit those 2 columns and other 2 variables which are irrelevant to the prediction.
  Total no. of variables: 19
  Number of data points: 10127
  
- *Type:*
  i) int (10 variables): CLIENTNUM, Customer_Age, Dependent_count, Total_Relationship_Count, Months_Inactive_12_mon, Contacts_Count_12_mon,
                       Total_Revolving_Bal, Total_Trans_Amt, Total_Trans_Ct
  ii) chr (6 variables): Attrition_Flag, Gender, Education_Level, Marital_Status, Income_Category, Card_Category
  iii) num (5 variables): Credit_Limit, Avg_Open_To_Buy, Total_Amt_Chng_Q4_Q1, Total_Ct_Chng_Q4_Q1, Avg_Utilization_Ratio
  
- *Omitting variables:*
  num (3 variables): 
  Naive_Bayes_Classifier_Attrition_Flag_Card_Category_Contacts_Count_12_mon_Dependent_count_Education_Level_Months_Inactive_12_mon_1, Naive_Bayes_Classifier_Attrition_Flag_Card_Category_Contacts_Count_12_mon_Dependent_count_Education_Level_Months_Inactive_12_mon_2, CLIENTNUM
  int (1 variable):
  Card_Category.1
  
- Binary Classification problem
  Target variable: Attrition_Flag
  No. of classes: 2 ("Existing Customer", "Attrited Customer")

#Data Exploration:

**Converting appropriate variables to factors**

```{r}
bank_churners$Attrition_Flag <- as.factor(bank_churners$Attrition_Flag)
bank_churners$Gender <- as.factor(bank_churners$Gender)
bank_churners$Education_Level <- as.factor(bank_churners$Education_Level)
bank_churners$Marital_Status <- as.factor(bank_churners$Marital_Status)
bank_churners$Income_Category <- as.factor(bank_churners$Income_Category)
bank_churners$Card_Category <- as.factor(bank_churners$Card_Category)
str(bank_churners)
```

##Summary statistics:

**5 number summary**

```{r}
summary(bank_churners)
```
Totally 1627 customers have churned the credit card service
There are no missing values and will work with outliers using boxplot while cleaning data.

##Visualizations:

**i) Understanding the distribution of each variables**

```{r, echo=TRUE}
library(ggplot2)
ggplot(bank_churners, aes(Customer_Age)) + geom_histogram()
```
There are some outliers above age 65.That might be because very few senior citizens have been using credit card.

```{r, echo =TRUE}
ggplot(bank_churners, aes(Dependent_count)) + geom_histogram()
```
Majority of the clients seemsto have 2 or 3 dependents

```{r, echo=TRUE}
ggplot(bank_churners, aes(Total_Relationship_Count)) + geom_histogram()
```
Most of the customers have 3 products from the bank

```{r, echo=TRUE}
ggplot(bank_churners, aes(Months_Inactive_12_mon)) + geom_histogram()
```
Most of the customers being inactive for 3 months

```{r, echo=TRUE}
ggplot(bank_churners, aes(Contacts_Count_12_mon)) + geom_histogram()
```

```{r, echo=TRUE}
ggplot(bank_churners, aes(Credit_Limit)) + geom_histogram()
```
Credit Limit variable is right skewed

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Total_Revolving_Bal)) + geom_histogram()
```
Most of the clients have 0 Revolving Balance

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Avg_Open_To_Buy)) + geom_histogram()
```
Again Avg_Open_To_Buy is having right skewed distribution

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Total_Amt_Chng_Q4_Q1)) + geom_histogram()
```
There are some outliers over here

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Total_Trans_Amt)) + geom_histogram()
```
This distribution is little weird

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Total_Trans_Ct)) + geom_histogram()
```
Here, I can see bimodal distribution

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Total_Ct_Chng_Q4_Q1)) + geom_histogram()
```
Again some outliers over here

```{r, echo=TRUE}
ggplot(bank_churners, aes(x=Avg_Utilization_Ratio)) + geom_histogram()
```
Right skewed distribution

**ii) Getting insights**

```{r, echo = TRUE}
p <- ggplot(bank_churners, aes(x=Months_Inactive_12_mon, fill=Attrition_Flag))
p + geom_bar(position="stack")
```
Most the customers have churned after being inactive for 2 or 3 months. 

```{r, echo = TRUE}
af <- ggplot(bank_churners, aes(x=Card_Category, fill=Attrition_Flag))
af + geom_bar(position="stack")
```
The clients who held the blue card have churned the most.

```{r, echo = TRUE}
q <- ggplot(bank_churners, aes(x=Total_Revolving_Bal, fill=Attrition_Flag))
q + geom_bar(position="stack")
```
About 900 customers with 0 revolving balance have quitted the credit card service

```{r, echo = TRUE}
s <- ggplot(bank_churners, aes(x=Income_Category, fill=Attrition_Flag))
s + geom_bar(position="stack")
```
```{r}
unique(bank_churners$Income_Category)
```

```{r, echo=TRUE}
r <- ggplot(bank_churners, aes(x=Income_Category, y=Credit_Limit, fill=Attrition_Flag))
r + geom_col()
```
People with income category $80k - $120k are given more credit limit and thereby we can see more existing customers in this category but still some of them have churned the credit card service.

**iii) Understanding relationship between variables**

```{r}
# Integers and numeric variables
bint <- bank_churners %>% select(c(2,4,9,10,11,12,13,14,15,16,17,18,19))

# Randomly sampling 1000 data to visualize scatterplot
library(dplyr)
sam <- sample_n(bint, 1000)
plot(sam)
```
From the above plot, I've seen following variable relationships:
-> Credit_Limit and Avg_Open_To_Buy
-> Credit_Limit and Avg_Utilization_Ratio
-> Avg_Open_To_Buy and Avg_Utilization_Ratio
-> Total_Trans_Amt, Total_Trans_Ct

*Performing scatterplots to find the relationships between variables*
```{r, echo=TRUE}
ggplot(bint, aes(Credit_Limit, Avg_Open_To_Buy)) + geom_point()
```
```{r, echo=TRUE}
ggplot(bint, aes(Credit_Limit, Avg_Utilization_Ratio)) + geom_point()
```
```{r, echo=TRUE}
ggplot(bint, aes(Avg_Open_To_Buy, Avg_Utilization_Ratio)) + geom_point()
```
```{r, echo=TRUE}
ggplot(bint, aes(Total_Trans_Amt, Total_Trans_Ct)) + geom_point()
```
*Pearson correlation*
```{r}
cor(bint)
```
*Strongly Correlated variables:*
-> Credit_Limit and Avg_Open_To_Buy => 0.995980544 (99.59% positively correlated)
-> Total_Trans_Amt, Total_Trans_Ct => 0.80719203 (80.71% positively correlated)
-> Total_Revolving_Bal and Avg_Utilization_Ratio => 0.624021991 (62.4% positively correlated)
-> Credit_Limit and Avg_Utilization_Ratio => -0.482965071 (48.29% negatively correlated)
-> Avg_Open_To_Buy and Avg_Utilization_Ratio => -0.538807748 (53.88% negatively correlated)

*Cross tabulation to find what kind of groups to concentrate in order to avoid further churn*

```{r, echo=TRUE}
xt1 <- bank_churners %>% group_by(Card_Category) %>% select(Card_Category, Attrition_Flag) %>% table()
xt1
```
1519 customers owning blue card have been churned.

```{r, echo=TRUE}
xt2 <- bank_churners %>% group_by(Customer_Age) %>% select(Customer_Age, Attrition_Flag) %>% table()
xt2
```
From the data, age group of 40-55 is more likely to quit the credit card service

```{r, echo=TRUE}
xt3 <- bank_churners %>% group_by(Income_Category) %>% select(Income_Category, Attrition_Flag) %>% table()
xt3
```
People with income less than $40k had attrited the most.

```{r, echo=TRUE}
xt4 <- bank_churners %>% group_by(Gender) %>% select(Gender, Attrition_Flag) %>% table()
xt4
```
930 females and 697 males have churned. Since both the numbers are high. Have to concentrate on both the groups.

```{r, echo=TRUE}
xt5 <- bank_churners %>% group_by(Education_Level) %>% select(Education_Level, Attrition_Flag) %>% table()
xt5
```

```{r, echo=TRUE}
xt6 <- bank_churners %>% group_by(Marital_Status) %>% select(Marital_Status, Attrition_Flag) %>% table()
xt6
```
I don't think education level and Marital status would affect the churning

#Data Cleaning

**Checking for outliers:**

```{r, echo=TRUE}
boxplot(bank_churners$Avg_Utilization_Ratio)
```
```{r, echo = TRUE}
boxplot(bank_churners$Total_Ct_Chng_Q4_Q1)
```
```{r, echo = TRUE}
boxplot(bank_churners$Total_Trans_Ct)
```
```{r, echo = TRUE}
boxplot(bank_churners$Total_Trans_Amt)
```
```{r, echo = TRUE}
boxplot(bank_churners$Total_Amt_Chng_Q4_Q1)
```
```{r, echo = TRUE}
boxplot(bank_churners$Avg_Open_To_Buy)
```
```{r, echo = TRUE}
boxplot(bank_churners$Total_Revolving_Bal)
```
```{r, echo = TRUE}
boxplot(bank_churners$Credit_Limit)
```
```{r, echo = TRUE}
boxplot(bank_churners$Contacts_Count_12_mon)
```
```{r, echo = TRUE}
boxplot(bank_churners$Total_Relationship_Count)
```
```{r, echo = TRUE}
boxplot(bank_churners$Dependent_count)
```
```{r, echo = TRUE}
boxplot(bank_churners$Customer_Age)
```

**Removing Outliers**

*Total_Ct_Chng_Q4_Q1*

```{r, echo=TRUE}
new <- bank_churners
new1 <- new %>% filter(Total_Ct_Chng_Q4_Q1 > 0.25 & Total_Ct_Chng_Q4_Q1 < 1.14)
boxplot(new1$Total_Ct_Chng_Q4_Q1)
```

*Total_Trans_Ct*

```{r, echo=TRUE}
new2 <- new1 %>% filter(Total_Trans_Ct < 134)
boxplot(new2$Total_Trans_Ct)
```
*Total_Trans_Amt*

```{r, echo=TRUE}
new3 <- new2 %>% filter(Total_Trans_Amt < 8087)
boxplot(new3$Total_Trans_Amt)
```
*Total_Amt_Chng_Q4_Q1*

```{r, echo=TRUE}
new4 <- new3 %>% filter( Total_Amt_Chng_Q4_Q1 < 1.16 & Total_Amt_Chng_Q4_Q1 > 0.28)
boxplot(new4$Total_Amt_Chng_Q4_Q1)
```
*Avg_Open_To_Buy*

```{r, echo=TRUE}
new5 <- new4 %>% filter(Avg_Open_To_Buy < 6965)
boxplot(new5$Avg_Open_To_Buy)
```
*Credit_Limit*

```{r, echo=TRUE}
new6 <- new5 %>% filter(Credit_Limit < 6500)
boxplot(new6$Credit_Limit)
```
*Contacts_Count_12_mon*

```{r, echo=TRUE}
new7 <- new6 %>% filter(Contacts_Count_12_mon < 5 & Contacts_Count_12_mon > 0.1)
boxplot(new7$Contacts_Count_12_mon)
```
*Months_Inactive_12_mon*

```{r, echo = TRUE}
new8 <- new7 %>% filter(Months_Inactive_12_mon < 5 & Months_Inactive_12_mon > 0.1)
boxplot(new8$Months_Inactive_12_mon)
```
*Customer_Age*


```{r, echo=TRUE}
new9 <- new8 %>% filter(Customer_Age<70)
boxplot(new9$Customer_Age)
```

**Dataset with no outliers**
```{r}
str(new9)
```
No. of features: 19
No. of instances: 4830

```{r}
summary(new9)
```

#Data Preprocessing

##Scaling
```{r}
library(caret)

scaling <- preProcess(new9, method=c("center", "scale"))
std_churn <- predict(scaling, new9)
summary(std_churn)
```

##Feature Selection

**Converting categorical variables to numerical variable using dummyVars function**

```{r}
dummy <- dummyVars(Attrition_Flag ~ ., data = std_churn)
dummies <- as.data.frame(predict(dummy, newdata = std_churn))
head(dummies)
```

**Removing near zero variance predictors**

```{r}
nzv <- nearZeroVar(dummies)
nzv
```
*Checking zero variance columns*

```{r}
names(dummies)
```

The following are zero variance columns:
6 - "Education_Level.Doctorate"
16 - "Income_Category.$120K +" 
22 -  "Card_Category.Blue"  
23 - "Card_Category.Gold"   
24 - "Card_Category.Platinum" 
25 -  "Card_Category.Silver"

```{r}
#Removing card category
bc_dummy <- dummies %>% select(-c(22,23,24,25))
nzv1 <- nearZeroVar(bc_dummy)
length(nzv1)
```

**Performing PCA**

```{r}
churn.pca <- prcomp(bc_dummy)
summary(churn.pca)
```

**Scree plot**

```{r, echo=TRUE}
screeplot(churn.pca, type = "l") + title(xlab = "PCs")
```
Here, the elbow is at 4. So, I'm choosing 4 principal components.

```{r}
# We don't want to include a prediction target variable in PCA, so we'll separate it

# Create the components
pre_pca <- preProcess(bc_dummy, method="pca", pcaComp=4)
churn.pc <- predict(pre_pca, bc_dummy)
# Put back target column
churn.pc$Attrition_Flag <- std_churn$Attrition_Flag
# Make sure that we have the PCs as predictors
head(churn.pc)

```
#Clustering

```{r}
# Removing label

pcwol <- churn.pc %>% select(-c(5))
head(pcwol)
```

##KMeans Clustering

**Finding k (no. of clusters)**

*Knee Plot*
```{r, echo=TRUE}
library(stats)
library(factoextra)

fviz_nbclust(pcwol, kmeans, method = "wss")
```
From knee plot, no. of clusters = 4

*Silhouette plot*

```{r, echo = TRUE}
fviz_nbclust(pcwol, kmeans, method = "silhouette")
```
From silhouette, no. of clusters = 3.

*I'm going to use 3 clusters for my clustering process*

**Fitting the data**

```{r}
fit <- kmeans(pcwol, centers = 3, nstart = 25)
fit
```

**Cluster Plot**

```{r, echo=TRUE}
fviz_cluster(fit, data = pcwol)
```

**Displaying PCA plot with clusters expressed in colors**

```{r, echo=TRUE}
pca = prcomp(bc_dummy)
rotated_data = as.data.frame(pca$x)
rotated_data$color <- std_churn$Attrition_Flag
ggplot(data = rotated_data, aes(x = PC1, y = PC2, col = color)) + geom_point(alpha = 0.3)
```
**Displaying plot with cluster results**

```{r, echo=TRUE}
# Assign clusters as a new column
rotated_data$Clusters = as.factor(fit$cluster)
# Plot and color by labels
ggplot(data = rotated_data, aes(x = PC1, y = PC2, col = Clusters)) + geom_point()
```
##Hierarchical Clustering

**Dissimilarity matrix**

```{r}
library(cluster)
dist_mat2 <- daisy(pcwol, metric = "gower")
# Result is a dissimilarity matrix
summary(dist_mat2)
```

**Knee plot**

```{r, echo=TRUE}
fviz_nbclust(pcwol, FUN = hcut, method = "wss")
```
From knee plot, no. of clusters for hierarchical clustering is 3

**Silhouette plot**

```{r, echo=TRUE}
fviz_nbclust(pcwol, FUN = hcut, method = "silhouette")
```
From silhouette plot, no. of clusters for hierarchical clustering is 3

**Building HAC Cluster model**

```{r}
hfit <- hclust(dist_mat2, method = 'complete')
h3 <- cutree(hfit, k=3)
```

**Visualizing HAC Cluster**

```{r, echo=TRUE}
fviz_cluster(list(data = pcwol, cluster = h3))
```

**Displaying cluster results with PCA projection**

```{r, echo=TRUE}
# Assign clusters as a new column
rotated_data$Clusters = as.factor(h3)
# Plot and color by labels
ggplot(data = rotated_data, aes(x = PC1, y = PC2, col = Clusters)) + geom_point()
```
**Displaying PCA plot with clusters expressed in colors** 

```{r, echo = TRUE}
ggplot(data = rotated_data, aes(x = PC1, y = PC2, col = color)) + geom_point()
```
**Comparing the results**

```{r, echo=TRUE}
# Create a dataframe
result <- data.frame(Status = std_churn$Attrition_Flag, HAC3 = h3, Kmeans = fit$cluster)
# View the first 20 cases one by one
head(result, n = 20)
```

*Clustering results In Cross-tabulation*

```{r}
# Crosstab for HAC
result %>% group_by(HAC3) %>% select(HAC3, Status) %>% table()
```

```{r}
# Crosstab for K Means
result %>% group_by(Kmeans) %>% select(Kmeans, Status) %>% table()
```
Here K Means seemed to get relatively better clusters on the data according to both cross tabulation and the one-by-one comparison on the table. HAC placed most of the data in the second cluster while struggling to cluster Attrited customers.

#Classification:

##SVM Classifier

Using 10 fold stratified cross validation

```{r}
folds <- 10

# Generate stratified indices
idx <- createFolds(churn.pc$Attrition_Flag, folds, returnTrain = T)

# Evaluation method
train_control_strat <- trainControl(index = idx, method = 'cv', number =
folds)

# Fit the model
svm_strat <- train(Attrition_Flag ~., data = churn.pc, method = "svmLinear", 
 trControl = train_control_strat)

# Evaluate fit
svm_strat
```

**Accuracy: 87.68%**

**Confusion Matrix**
```{r}
confusionMatrix(svm_strat, churn.pca$Attrition_Flag)

```

It has wrongly predicted 446 Attrited customers as Existing customers.

**Grid Search**

```{r}
grid <- expand.grid(C = seq(0.001, 2, length = 20))

# Fit the model
svm_stratcv_grid <- train(Attrition_Flag ~., data = churn.pc, method = "svmLinear", 
 trControl = train_control_strat, tuneGrid = grid)

# View grid search result
svm_stratcv_grid
```

Accuracy: 87.78%

##KNN Classifier

```{r}
set.seed(123)

# 10 fold cross-validation
ctrl <- trainControl(method="cv", number = 10) 
knnFit <- train(Attrition_Flag ~ ., data = churn.pc, 
 method = "knn", 
 trControl = ctrl)

#Output of kNN fit
knnFit
```

Accuracy: 89.04%

**Accuracy vs No. of Neighbors**
```{r}
knnFit1 <- train(Attrition_Flag ~ ., data = churn.pc, method = "knn", trControl = ctrl, tuneLength = 10)

plot(knnFit1)

```
Accuracy is high at k=9

**Accuracy of SVM: 87.78%**
**Accuracy of KNN: 89.04%**

#Evaluation:

Better classifier: KNN

**Generating Confusion Matrix for KNN**

```{r}
pred_knn<- predict (knnFit, churn.pc)
confusionMatrix(churn.pc$Attrition_Flag, pred_knn)
```
151 Attrited customers were wrongly predicted as Existing customers. From the confusion matrix, this error is critical.

From Confusion Matrix:

Positive class: Attrited Customers
Negative class: Existing Customers

TP: 517     FN: 151
FP: 287     TN: 3875


**Calculating Precision**

Precision = TP / (TP + FP)
          = 517 / (517 + 287)
          = 0.643

**Calculating Recall**

Recall = TP / (TP + FN)
       = 517 / (517 + 151)
       = 0.7739
       
**ROC Plot**

```{r}
#Target class
str(churn.pc$Attrition_Flag)
unique(churn.pc$Attrition_Flag)
```

```{r}
library(pROC)

# Get class probabilities for KNN
pred_prob <- predict(knnFit, churn.pc, type = "prob")
head(pred_prob)
```
```{r, echo=TRUE}
# And now we can create an ROC curve for our model.
roc_obj <- roc((churn.pc$Attrition_Flag), pred_prob[,1])
plot(roc_obj, print.auc=TRUE)
```

Accuracy: 90.93% 
Precision: 64.3%
64.3% of postive class was predicted correctly among all positives
Recall: 77.39%
77.39% of actual positives were correctly predicted as positive
AUC: 0.952 
This value seems good but it is misleading as the data has class imbalance problem
Sensitivity : 77.4% (True Positive Rate)
Specificity : 93.10% (True Negative Rate)
Balanced Accuracy: 85.25%

#Report

**From data exploration:**
1.Most of the customers have churned after being inactive for 2 or 3 months.So, target on the people who are inactive for 2-3 months. As they're likely to be attrited
2.The clients who held the blue card have churned the most. So, it's recommended to change the benefits of the card making favourable to the customers.1519 customers owning blue card have been churned.
3.Most of the clients having 0 Revolving Balance are likely to be churned.About 900 customers with 0 revolving balance have quitted the credit card service
4.People with income category $80k - $120k are given more credit limit and thereby we can see more existing customers in this category but still some of them have churned the credit card service.
5. People with age group of 40-55 is more likely to quit the credit card service, so have to turn focus on them too

Approach to following groups of people to minimize the churning rate to some extent
Card category: Blue
Age: 40-55
Income : Less than $40K
Blue card holders

**From Clustering**
K Means seemed to get relatively better clusters on the data. HAC placed most of the data in the second cluster while struggling to cluster Attrited customers.

**From Classifiers**
KNN seems to be good as it has high accuracy but it is misleading due to the class imbalance problem in data as there were more existing customers than credit card service churned customers

**From Evaluation:**
Accuracy, Balnaced Accuracy and AUC values are good. Intolerable error from the confusion matrix is that 151 Attrited customers were wrongly predicted as Existing customers. 